%
% Specifying solver parameters via XML parameter file
%
\chapter{Solver Parameters}
%
%
\section{Interface between CFS++ and OLAS}
\label{sec:SP_Interface}
%
%
This section deals with specifying solvers and preconditioners for the linear
systems associated with the solution of the different PDEs. It is important to
note at this point that these parameters are of no importance in CFS++ itself,
but will only be passed on to the underlying linear algebra system OLAS. The
necessary interface in CFS++ is proviede by the CFSOLASParams class.

Note that all parameters that can be specified for the individual solvers and
preconditioners are optional. Contrary to the policy in other cases,
default values for missing parameters are currently not specified via the XML
Schema definition and the XMLParamHandler class, but are prescribed in the
SetDefaultParams() methods of the OLAS\_Params class, which constitutes the
method for shuffling parameters from CFS++ to OLAS. See that file or the
doxygen documentation of OLAS for the defaults.
%
%
\section{Basic specification}
%
%
The PDE section may contain a \xml{solver} element. The element is optional
and can be used to choose a solution algorithm for the linear system
associated to the PDE and in the case of iterative solvers also a
preconditioning method. The two choices are specified as the \xml{type}
and \xml{precond} attributes of the \xml{solver} element. The following two
tables give an overview on the currently available choices for solver and
preconditioner:
%
%
\begin{center}
\begin{tabular}{lp{0.65\textwidth}}
\hline\noalign{\smallskip}
\multicolumn{2}{c}{Currently available choices for \xml{type} attribute of
\xml{solver} element}\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Attribute value & solver\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Richardson    & richardson iteration\\
cg	      & (preconditioned) conjugate gradient algorithm\\
gmres	      & generalised minimal residual method\\
QMR	      & quasi-minimal residual method\\
hyprePCG      & hypre's version of the classical conjugate gradient algorithm\\
hypreGMRES    & hypre's version of the generalised minimal residual method\\
hypreBICGSTAB & hypre's version of the stabilised bi-conjuigate gradient
                method\\
\noalign{\smallskip}\hline\noalign{\smallskip}
directLU      & direct solution via $LU$ decomposition\\
directLDL     & direct solution via $LDL^T$ decomposition (planned)\\
lapackLU      & direct solution via LAPACK\\
lapackLL      & direct solution via LAPACK (for symmetric/hermitean, positive
                definite matrices only)\\
pardiso       & direct solution with sparse direct solver from Pardiso
                project\\
\noalign{\smallskip}\hline\noalign{\smallskip}
expertsChoice & let the Expert() module decide which solver to use
                \emph{(default value, if nothing is specified)}\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\begin{center}
\begin{tabular}{lp{0.65\textwidth}}
\hline\noalign{\smallskip}
\multicolumn{2}{c}{Currently available choices for precond attribute of solver
element}\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Attribute value & preconditioner\\
\noalign{\smallskip}\hline\noalign{\smallskip}
noPrecond   & no preconditioning (for direct solvers)\\
Id	    & preconditioning with identity matrix, i.e. no preconditioning
              (for iterative solvers)\\
MG	    & proprietory algebraic multigrid method of OLAS\\
Jacobi	    & plain Jacobi preconditioning\\
SSOR	    & symmetric successive over-relaxation\\
ILU	    & incomplete LU decomposition using ILU(0)\\
hypreJacobi & plain Jacobi in hypre\\
hypreAMG    & hypre's algebraic multigrid method\\
hypreILU    & hypre's ILU preconditioner Euclid, ILU(k) currently\\
hypreSPAI   & hypre's SPAI preconditioner ParaSails \\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
The default values for \xml{type} and \xml{precond} are \xml{expertsChoice} and
\xml{noPrecond}.
%
%

The \xml{solver} element can have several children which can be used to specify
e.g.~properties of the system matrix and parameters for the chosen solver and
preconditioner. Its general structure is as follows:
%
%
\begin{center}
\begin{minipage}{0.6\textwidth}
\begin{tabbing}
\quad\=\\ \kill
\verb!<solver type="!\emph{value}\verb!" precond="!\emph{value}\verb!">!\\
\> \verb!<matrix eqnNumbering="!\emph{value}\verb!" entry="!\emph{value}\verb!" storage="!\emph{value}\verb!"/>!\\
\> \verb!<exportLinSys baseName="!\emph{value}\verb!"/>!\\
\> \verb!<stoppingRule type="!\emph{value}\verb!"/>!\\
\> \emph{element with parameters for solver}\\
\> \emph{element with parameters for preconditioner}\\
\verb!</solver>!\\
\end{tabbing}
\end{minipage}
\end{center}
%
%
\section{Specifying Properties of the System Matrix}
%
%
The first child of the \xml{solver} element can be a \xml{matrix} element. The
matrix element possesses the following four optional attributes:
\xml{eqnNumbering}, \xml{entry}, \xml{storage} and \xml{reordering}.
%
%
\begin{center}
\begin{longtable}{llp{0.5\textwidth}}
\hline\noalign{\smallskip}
attribute & values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
eqnNumbering & scalar & Every unknown receives a unique equation number.
                        Unknowns whose value is fixed by homogeneous Dirichlet
                        boundary conditions are eliminated. Unknowns given by
                        inhomogeneous Dirichlet boundary conditions are treated
                        with the penalty approach.

                        In this case the matrix representing the linear system
                        will have scalar double or complex entries.\\
             & block  & This is relevant only for PDEs like e.g.~mechanic,
                        where to each node of the grid belong several unkowns.
                        If block is chosen as value, then all unknowns
                        belonging to a node, so actually the node itself, will
                        receive a unique equation number. If all the unknowns
                        at a node are fixed by homogeneous Dirichlet boundary
                        conditions, then the node and all its unknowns are
                        eliminated. If not all unkowns are fixed, or if
                        inhomogeneous Dirichlet boundary conditions are
                        involved, then the penalty approach is employed.

                        In this case the matrix representing the linear system
                        will have square matrices of dimension $n\times n$ as
                        entries, where $n$ is the number of degrees of freedom
                        per node. The tiny matrices themselves will have scalar
                        double or complex entries.\\
             & superBlock & This setting is currently only of importance for
                        the piezo PDE. It will be no longer required, as soon
                        as direct coupling is implemented in CFS++.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
entry & double, complex & specifies the entry type on the scalar level, i.e.
                          the type of the entries of the system matrix for
                          scalar equation numbering, or the type of the entries
                          of the tiny matrices in block numbering.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
storage & sparseSym      & System matrix is stored in compressed row storage
                           format (CRS) format. However, symmetry is exploited
                           by storing only the upper triangle.\\
        & sparseNonSym	 & System matrix is stored in CRS format.\\
        & skylineSym	 & \emph{not fully supported yet}\\
        & skylineNonSym	 & \emph{not fully supported yet}\\
        & hypreMatrix	 & System matrix is stored as a hypre matrix. This is
	                   the correct choice, if a hypre solver is to be used.
                           Note that a hypre matrix can only store scalar
                           double entries.\\
        & lapackGBMatrix & System matrix is stored as a general banded matrix.
                           This is the suitable choice for the lapackLU solver.
                           \\
        & lapackPBMatrix & \emph{not fully supported yet}\\
        & expertsChoice  & CFS++ will choose the storage type itself.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
reordering & noReordering & Specifies that OLAS should not reorder the graph.\\
           & Sloan        & Specifies that OLAS should try to reorder the
                            graph, before the associated matrices are
                            assembled, using the bandwidth reduction algorithm
                            by Sloan. Note that the reordering will be
                            discarded, if the profile of the matrix is not
                            reduced.\\
           & Metis        & Specifies that OLAS should try to reorder the
                            graph, before the associated matrices are
                            assembled, using a fill-reducing ordering from
                            Metis. This will be a sort of nested-discetion
                            reordering. Note that this is untested so far.\\
\noalign{\smallskip}\hline
\end{longtable}
\end{center}
%
%
Note that not all values of \xml{eqnNumbering}, \xml{entry} and \xml{storage}
will work together and that some choices, while being possible will not lead
to an efficient simulation run in combination with certain solvers. The
CFSOLASParams interface class has an Expert() subroutine, that will adapt the
parameters, if they are not sensible and issue a respective warning. It is also
this routine that will determine a storage type for the system matrix, if
\xml{expertsChoice} is specified, which also is the default for the
\xml{storage} attribute.
%
%
\section{Exporting the Linear System}
%
%
In some situations one is interested in obtaining the system matrix of the
linear system for post-processing. We may e.g.~want to plot the sparsity
pattern or compute some eigenvalues using matlab. Thus, OLAS offers the
possibility to export the system matrix to an ascii file in a format conforming
to the MatrixMarket specifications. For details see
\url{math.nist.gov/MatrixMarket}. The right-hand side vector is exported
together with the matrix, but in a separate ascii file. The first line of this
file contains the dimension of the vector. This is followed by the vector
entries.

In order to trigger the export one must specify an \xml{exportLinSys} element.
This is a child of the \xml{solver} element. It expects as \xml{baseName}
attribute the basename for the two output files. For the MatrixMarket file a
trailing '.mtx' will be appended to this basename, while for the right-hand
side a '.vec' postfix is used. See Sec.~\ref{sec:SP_Examples} for an example.
%
%
\section{Setting the stopping rule}
%
%
OLAS supports three different kinds of stopping rules for iterative solvers.
One chooses a certain criterion by setting
%
%
\begin{center}
\verb!<stoppingRule type="!\emph{value}\verb!"/>!
\end{center}
%
%
where \emph{value} must be one of the possibilities from the list below:
%
%
\begin{center}
\begin{tabular}{lc}
\hline\noalign{\smallskip}
value & stopping test \\
\noalign{\smallskip}\hline\noalign{\smallskip}
{absNorm}     & $\displaystyle\|r^{(k)}\|_2 \leqslant \varepsilon$\\
\noalign{\smallskip}\hline\noalign{\smallskip}
{relNormRHS}  & $\displaystyle\frac{\|r^{(k)}\|_2}{\|b\|_2} \leqslant
 \varepsilon$\\
\noalign{\smallskip}\hline\noalign{\smallskip}
{relNormRes0} & $\displaystyle\frac{\|r^{(k)}\|_2}{\|r^{(0)}\|_2}\leqslant
\varepsilon$\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
Here $b$ is the right-hand side of the linear system $Ax=b$ under consideration
and $r^{(k)}=b - Ax^{(k)}$ is the residual of the approximate solution
$x^{(k)}$ after $k$ iteration steps.

In some cases choosing \xml{relNormRHS} will lead to difficulties. One such
case is when $b=0$, another one when the penalty formulation is employed to
treat inhomogeneous Dirichlet boundary conditions. Both cases will be detected
by OLAS automatically and \xml{relNormRes0} will be used instead of
\xml{relNormRHS}.
%
%
\section{Parameters for special solvers}
%
%
This section lists for each solver the parameters that can be specified to
steer its behaviour. These parameters are given in the following fashion.
In order to specify parameters for solver \xml{xyz} the \xml{solver} element
is supplemented with a child element with name \xml{xyz}. The parameters
are then given as children of this sub-element. The order of these child
elements is reflected by the ordering of the parameters in the respective
tables. See Sec.~\ref{sec:SP_Examples} for an example.

All the solver parameters are optional as is the \xml{xyz} element. If one are
all parameters are omitted the respective default value defined within OLAS
will be used, cf.~Sec.~\ref{sec:SP_Interface}.
%
%
\subsection{Pardiso}
%
%
Pardiso is a (parallel) sparse direct solver developed mainly at the University
of Basel. We cite the web-page
(\url{http://www.computational.unibas.ch/cs/scicomp/software/pardiso}):
%
%
\begin{quote}
The package PARDISO is a thread-safe, high-performance, robust, memory efficient and easy to use software for solving large sparse symmetric and unsymmetric linear systems of equations on shared memory multiprocessors. The solver uses a combination of left- and right-looking Level-3 BLAS supernode techniques to exploit pipelining parallelism and memory hierarchies. In order to improve sequential and parallel sparse numerical factorization performance, the algorithms are based on a Level-3 BLAS update and pipelining parallelism is exploited with a combination of left- and right-looking supernode techniques.

For sufficiently large problem sizes numerical experiments demonstrate that the scalability of the parallel algorithm is nearly independent of the shared-memory multiprocessing architecture and speedup of up to seven using eight processors have been observed. The approach is based on OpenMP directives and has been successfully tested on the following parallel computers: COMPAQ AlphaServer, AMD Opteron, SGI Origin 2000, SUN Enterprise Server, Intel Pentium IV / Itanium(R) with RedHat LINUX, NEC SX-5, and Cray J90 and IBM Regatta SMPs.
\end{quote}
%
%
OLAS offers an interface to use Pardiso as solver for a linear system. The
interface accepts the following parameters:
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
posDef     & yes, no & Set to yes, if matrix is positive definite\\
\noalign{\smallskip}\hline\noalign{\smallskip}
hermitean  & yes, no & Set to yes, if matrix is hermitean\\
\noalign{\smallskip}\hline\noalign{\smallskip}
symStruct  & yes, no & Set to yes, if matrix has a symmetric pattern
                       of non-zero entries\\
\noalign{\smallskip}\hline\noalign{\smallskip}
ordering & nestedDissection & in this case Pardiso will use Metis (contained
                              in the Pardiso library) to compute a
                              fill-reducing re-ordering for the matrix.\\
         & minimumDegree    & for this choice Pardiso uses a minimum degree
                              algorithm for computing a
                              fill-reducing re-ordering for the matrix.\\
         & noReordering     & in this case Pardiso is tricked into using the
                              original matrix ordering determined by OLAS.
                              This is mainly intended for testing the orderings
                              implemented in OLAS, like e.g.~Sloan, in
                              comparison with those offered by Pardiso.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging    & yes, no & If set to yes, the solver object will write minimal
                       status reports to the standard OLAS logfile\\
\noalign{\smallskip}\hline\noalign{\smallskip}
stats      & yes, no & Setting this parameter to yes triggers Pardiso to print
                       statistical information to the standard output\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
Note:
\begin{itemize}
\item The interface is designed for version 1.2.2
\item Current Pardiso licensing policy requires a user to obtain a personal
      license for each (username, hostname) combination.
\item If you use Pardiso as solver, it usually makes no sense to set
      the \xml{reordering} attribute of the \xml{solver} element to some other
      value besides the default of \xml{noReordering}, unless you use
      \xml{ordering} = \xml{noReordering} in Pardiso.
\end{itemize}
%
%
\subsection{lapackLU / lapackLL}
%
%
LAPACK is the standard package in scientific computing for high-level problems
in (dense) linear algebra. We cite the web-page
(\url{http://www.netlib.org/lapack}):
%
%
\begin{quote}
LAPACK is a library of Fortran 77 subroutines for solving the most commonly
occurring problems in numerical linear algebra. It has been designed to be
efficient on a wide range of modern high-performance computers. The name LAPACK
is an acronym for Linear Algebra PACKage.
\end{quote}
%
%
LAPACK does not support operations on sparse matrices, but offers solution
routines for problems involving matrices with banded structure. Thus, it is
an interesting alternative for small sized problems (after a band reducing
re-ordering).

OLAS provides two interfaces two the band solvers in LAPACK. lapackLU and
lapackLL. These classes support scalar matrices with either real or complex
entries. The lapackLU class will compute an LU decomposition and can thus, be
used for any type of problem. The lapackLL class computes a Cholesky
decomposition $LL^T$ and is thus limited to symmetric/hermitean positive
definite problems. The lapackLU solver accepts the following parameters
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tryScaling & yes, no & If set to yes, then the solver will try to equilibrate
                       the matrix before solving the linear system, in order
                       to improve the precision of the solution.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
refineSol  & yes, no & If set to yes, then the solver will perform some steps
                       of iterative refinement in order to improve the
                       precision of the solution.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging    & yes, no & If set to yes, then the solver will log some information
                       to the standard OLAS log stream.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
The lapackLL solver currently only supports \xml{logging}. Please also note
the following:
%
%
\begin{itemize}
\item The Lapack\_LU solver class currently makes use of its own LapackGBMatrix
      matrix class. Since the MultAdd() method is not implemented in this class
      the solver cannot be used for transient problems.
\item Do not set \xml{tryScaling} to yes, if you have inhomogeneous Dirichlet
      boundary conditions, since it will mess with the penalty approach!
\end{itemize}
%
%
\subsection{directLU / directLDL}
%
%
The directLDL solver computes in the setup phase a decomposition of the problem
matrix $A$ in the form $A=LDL^T$, with a lower triangular matrix $L$ and a
diagonal matrix $D$. For symmetric, positive definite matrices this corresponds
to a Cholesky decomposition $\hat{L}\hat{L}^T$, but with out the drawing of the
square root, i.e. $\hat{L} = LD^{-1/2}$. For symmetric indefinite matrices a
Cholesky decomposition does not exist, and only an $LDL^T$ factorisation can be
computed. In the complex case the situation is similar.\\[2ex]
%
%
By default the directLDL solver does not need any parameters
Parameters for the directLDL solver are:
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
saveFacFile     & string  & this parameter can be used to specify the name of
	                    a file to which the factor matrix $F=D+L^T$ is
                            exported using the MatrixMarket format
                            specification; if not specified, the matrix is not
                            exported\\
savePatternOnly & yes, no & if set to \xml{yes}, then only the sparsity pattern
                            of the matrix is exported, but not the numerical
                            values of the factorisation\\
logging         & yes, no & if set to \xml{yes}, the solver will be verbose,
                            i.e. it will write information on its analysis and
                            factorisation phase and on the individual solves
                            to the standard logfile.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{CG}
Setting \xml{type} to \xml{cg} selects the native preconditioned conjugate
gradient method implemented of OLAS.
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tol     & $> 0$        & threshold for stopping criterion\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxIter & $\mathbb{N}$ & maximal number of iterations\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging & yes, no & If logging is set to yes, then the solution process will
	            log the convergence history (absolute and relative residual
                    norm) to the standard OLAS log stream.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{GMRES}
%
%
Setting \xml{type} to \xml{gmres} selects the OLAS implementation of the
(re-started) Generalised Minimal Residual solver. This is a straightforward
implementation using the modified Gram-Schmidt process for computation of the
basis of the underlying Krylov subspace and Givens rotations for the solution
of the related least-squares problem. It can deal with real-valued as well as
complex-valued matrices. Its behaviour can be steered using the following
parameters
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tol & $> 0$ & The value of this parameter is used for the convergence criterion
              in the stopping test of GMRES. The stopping rule is as follows
	\begin{displaymath}
        \|r^{(k)}\|_2  \leq \varepsilon \cdot \|b\|_2
	\end{displaymath}
        where $r^{(k)}$ is the residual of the k-th step and $b$ is the right
        hand side. If $\|b\|_2=0$, then it is replaced by $\|r^{(0)}\|_2$.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxIter & $\mathbb{N}$ & Maximal number of iteration steps allowed for the
                         GMRES(m) method. If maxIter is set to 1 we actually
	                 perform a full GMRES iteration and the maximal
	                 number of iterations is limited by maxKrylovDim.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxKrylovDim & $\mathbb{N}$ & This determines the upper limit on the size of
	                      the Krylov subspace generated by the GMRES method
                              before it is re-started.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging & yes, no & If logging is set to yes, then the solution process will
	            log the convergence history (absolute and relative residual
                    norm) to the standard OLAS log stream.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{hypreGMRES}
%
%
Setting \xml{type} to \xml{hypreGMRES} selects the Generalised Minimal Residual
solver from the hypre library. Its behaviour can be steered using the following
parameters
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tol & $> 0$ & The value of this parameter is used for the convergence criterion
              in the stopping test of GMRES. The stopping rule is as follows
	\begin{displaymath}
        \|r^{(k)}\|  \leq \varepsilon \cdot \|b\|
	\end{displaymath}
        where $r^{(k)}$ is the residual of the k-th step and $b$ is the right
        hand side. If $\|b\|=0$, then it is replaced by $\|r^{(0)}\|$.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxIter & $\mathbb{N}$ & Maximal number of iteration steps allowed for the
                         GMRES method.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxKrylovDim & $\mathbb{N}$ & This gives the upper limit on the size of the
                              Krylov subspace generated by the GMRES method
                              before it is re-started. At least that's what
                              Marcus assumes from reading hypre's minimal
                              doc.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging & 0, 1, 2 & This parameter is used to control the verbosity of the
                    method. Its values have the following meaning:
\begin{center}
\begin{tabular}{lcl}
0 &=& no printout, no storing of norms\\
1 &=& no printout, norms are stored\\
2 &=& norms are stored and reported to\\
  & & stdout in every iteration step\\
\end{tabular}
\end{center}
It appears that this concept is new and not quite working yet.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
printLevel & 0, 1, 2 & This seems to be the old way of setting the verbosity.
                       See description of logging above.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{hypreBICGSTAB}
%
%
Setting \xml{type} to \xml{hypreBICGSTAB} selects the stabilised bi-conjugate
gradient solver from the hypre library. Its behaviour can be steered using the
following parameters
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tol & $> 0$ & The value of this parameter is used for the convergence criterion
              in the stopping test of GMRES. The stopping rule is as follows
	\begin{displaymath}
        \|r^{(k)}\|  \leq \varepsilon \cdot \|b\|
	\end{displaymath}
        where $r^{(k)}$ is the residual of the k-th step and $b$ is the right
        hand side. If $\|b\|=0$, then it is replaced by $\|r^{(0)}\|$.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxIter & $\mathbb{N}$ & Maximal number of iteration steps allowed for the
                         GMRES method.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging & 0, 1, 2 & This parameter is used to control the verbosity of the
                    method. Its values have the following meaning:
\begin{center}
\begin{tabular}{lcl}
0 &=& no printout, no storing of norms\\
1 &=& no printout, norms are stored\\
2 &=& norms are stored and reported to\\
  & & stdout in every iteration step\\
\end{tabular}
\end{center}
It appears that this concept is new and not quite working yet.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
printLevel & 0, 1, 2 & This seems to be the old way of setting the verbosity.
                       See description of logging above.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{hyprePCG}
%
%
Setting \xml{type} to \xml{hyprePCG} selects the preconditioned conjugate
gradient solver from the hypre library. Its behaviour can be steered using the
following parameters
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tol & $> 0$ & The value of this parameter is used for the convergence criterion
              in the stopping test of GMRES. The stopping rule is as follows
	\begin{displaymath}
        \|r^{(k)}\|  \leq \varepsilon \cdot \|b\|
	\end{displaymath}
        where $r^{(k)}$ is the residual of the k-th step and $b$ is the right
        hand side. If $\|b\|=0$, then it is replaced by $\|r^{(0)}\|$.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxIter & $\mathbb{N}$ & Maximal number of iteration steps allowed for the
                         GMRES method.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
logging & 0, 1, 2 & This parameter is used to control the verbosity of the
                    method. Its values have the following meaning:
\begin{center}
\begin{tabular}{lcl}
0 &=& no printout, no storing of norms\\
1 &=& no printout, norms are stored\\
2 &=& norms are stored and reported to\\
  & & stdout in every iteration step\\
\end{tabular}
\end{center}
It appears that this concept is new and not quite working yet.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
printLevel & 0, 1, 2 & This seems to be the old way of setting the verbosity.
                       See description of logging above.\\
\noalign{\smallskip}\hline\noalign{\smallskip}
twoNorm & yes, no & If this is set to yes, then the Euclidean norm is used
	            in the stopping test of CG instead of the energy norm
	            induced by the system matrix.\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{lapackLU}
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
tryScaling & yes, no & if set to yes, we let LAPACK try to improve the
                       condition number of the linear system by scaling the
                       rows and columns of the matrix before computing the
                       decomposition\\
refineSol  & yes, no & if set to yes, we let LAPACK perform some steps of
                       iterative refinement after the foward/backward
                       substitution\\
logging    & yes, no & if set to yes, LAPACK will be verbose about its solution
	               process, i.e. it will write information to the standard
	               logfile\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\section{Parameters for special preconditioners}
%
%
\subsection{MG}
%
%
\begin{center}
\begin{longtable}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \endhead
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
maxCoarseDepend & $\mathbb{N}$ &
The maximal number of coarse neighbours, an F-point can get due to their
strong influence on that point. This does not mean, that this will be the
number of C-neighbours, that will maximally surround an F-point, but the number
of C-neighbours, that will cause an F-point to be judged as
\emph{well interpolated}. Thus, a further neighbour will not be made a
C-point due to its dependency to the named point, but it could still get
C-point because of other reasons.\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
minSystemSize & $\mathbb{N}$ &
The minimal size for the coarsest system. If the coarsening process yields a
number of C-points for the next coarser system that falls below this limit, the
coarsening result is rejected and the corresponding hierarchy level is not
created. The present level gets the coarsest.\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
numPreSmooth & $\mathbb{N}_0$ &
Number of pre-smoothing steps in the AMG solution cycle\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
numPostSmooth & $\mathbb{N}_0$ &
Number of post-smoothing steps in the AMG solution cycle\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
cycleParameter & $\mathbb{N}$ &
Solution phase. The recursion parameter for the solution cycle, which is the
number of coarse corrections, applied \emph{on each level except the top
and the bottom level}. A value of 1 for example causes a common V-cycle, 2 a
common W-cycle. Note that excluding the top level from this multiple coarse
grid corrections causes the parameter value 2 to generate a real W-cycle, not
a WW-cycle.\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
alpha & $[0,1]$ &
The ``alpha'' in the classical Ruge-St\"{u}ben setup phase. It controls the
decision whether a dependency is strong or weak:
$(i,j) \in S \Leftrightarrow d(i,j) \ge \alpha$, with the dependency strength
of point $i$ from point $j$
$\displaystyle d(i,j) := \frac{|a_{ij}|}{\max\{|a_{ik}| : k \ne i\}}$ and
matrix entries $a_{ij}$.\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
strongDiagRatio & $[0,1]$ &
Additional necessary condition for a dependency to be judged as strong. The
alpha is related to the size of an off-diagonal entry compared to the other
off-diagonal entries. But it must be also be large in comparison with the
corresponding diagonal entry. strongDiagRatio represents this ratio:
$(i,j) \in S \Rightarrow \mbox{strongDiagRatio} \cdot a_{ii}) \le |a_{ij}|$.\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
forceFineRatio & $> 0$ &
If a row of the system matrix is strongly diagonal dominant, it can be forced
to get an F-point. AMG\_\-Force\-Fine\-Ratio is the limit value for the ratio
$ \frac{\sum_j |a_{ij}|}{a_{ii}} $. If this ratio falls below forceFineRatio,
the point is forced immediately to get an F-point. This is in particular useful
 for example in systems arrising from a PDE discretization in which Dirichlet
boundary conditions are treated using penalty terms. In this case all Dirichlet
points are forced to get F-points.\\
%
%
\noalign{\smallskip}\hline\noalign{\smallskip}
%
%
directSolver & noSolver $|$ lapackLU &
Selects the direct solver that will be used to solve the system on the coarsest
level. In the case that \xml{noSolver} is chosen the smoother will be used to
solve the problem.\\
\noalign{\smallskip}\hline
\end{longtable}
\end{center}
%
%
\subsection{hypreAMG}
%
%
Setting \xml{precond} to \xml{hypreAMG} selects the BoomerAMG preconditioner
from the hypre library. BoomerAMG is an algebraic multigrid method. The
following table lists the parameters currently supported by the OLAS--hypre
interface. For a detailed description of BoomerAMG see the hypre documentation.
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
maxLevels & $\mathbb{N}$     & maximal number of levels in multigrid
                               hierarchy\\
alpha     & [0,1]            & central threshold parameter for judging strength
                               of node couplings\\
numSweeps & $\mathbb{N}$     & number of multigrid cycles to perform in each
                               preconditioning step\\
cycleType & V-cycle, W-cycle & specifies the cycling strategy\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{hypreILU}
Setting \xml{precond} to \xml{hypreILU} selects the Euclid preconditioner
from the hypre library. Euclid is an incomplete LU decomposition
preconditioner. Currently it only officially supports ILU(k) preconditioning.
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
level  & $\mathbb{N}_0$ & generation level of allowed fill-in for ILU(k)
                          approach\\
stats  & yes, no        & print some statistics to standard output\\
memory & yes, no        & print memory some statistics to standard output\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{hypreSPAI}
%
%
Setting \xml{precond} to \xml{hypreSPAI} selects the ParaSails preconditioner
from the hypre library. ParaSails can be used as a SPAI or in the symmetric
positive definite case also as FSAI preconditioner. It uses a pre-selected
sparsity pattern for the approximate inverse, which it derives by a powers of
sparsified matrix (PSM) approach.
%
%
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
thresh      & $[-1,1]$     & threshold for determining strong connections in
                             the PSM approach\\
levels      & $\mathbb{N}$ & number of levels for the PSM approach\\
symmetry    & 0, 1, 2      & chooses between SPAI and FSAI\\
loadBalance & yes, no      & only important in parallel context\\
reuse       & yes, no      & do we want to re-use a sparsity pattern\\
\noalign{\smallskip}\hline
\end{tabular}
\end{center}
%
%
\subsection{hypreJacobi}
%
%
This is a simple Jacobi preconditioner. Thus, it does not require or allow the
specification of any parameters.
%
%
\subsection{Incomplete LDL preconditioners}
%
%
Incomplete LDL preconditioners work with symmetric matrices and compute a
decomposition of the problem matrix $A$ in the form
%
%
\begin{displaymath}
A = LDL^T - R
\enspace,
\end{displaymath}
%
%
where $L$ is a lower triangular matrix with ones on the diagonal, $D$ is a
diagonal matrix and $R$ is a remainder matrix, since the factorisation is
incomplete and $A\neq LDL^T$. OLAS currently supports three different version
of ILDL preconditioners: ILDL$(k)$, ILDL$(\tau,p)$ and ILDL$_{cn}$. All three
preconditioners share the following three parameters:
%
%
\begin{center}
\begin{longtable}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
element & range of values & meaning \endhead
\noalign{\smallskip}\hline\noalign{\smallskip}
saveFacFile     & string  & this parameter can be used to specify the name of
	                    a file to which the factor matrix $F=D+L^T$ is
                            exported using the MatrixMarket format
                            specification; if not specified, the matrix is not
                            exported\\
savePatternOnly & yes, no & if set to \xml{yes}, then only the sparsity pattern
                            of the matrix is exported, but not the numerical
                            values of the factorisation\\
logging         & yes, no & if set to \xml{yes}, the preconditioners will be
                            verbose, i.e. they will write information on their
                            setup and application to the standard logfile.\\
\noalign{\smallskip}\hline
\end{longtable}
\end{center}
%
%
Besides these common parameters each preconditioner allows to specify its own
set of steering parameters, which must be speficied in front of the common
ones. In the case of the ILDL$(\tau,p)$ preconditioner there are two specific
parameters:
\begin{center}
\begin{tabular}{lcp{0.5\textwidth}}
\hline\noalign{\smallskip}
\multicolumn{3}{c}{specific parameters for ILDL$(\tau,p)$}\\
\noalign{\smallskip}\hline\noalign{\smallskip}
element & range of values & meaning \\
\noalign{\smallskip}\hline\noalign{\smallskip}
fillVal   & $\mathbb{N}$ & \\
threshold & $[0,1]$      &  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\multicolumn{3}{c}{specific parameters for ILDL$(k)$}\\
\noalign{\smallskip}\hline\noalign{\smallskip}
level     & $\mathbb{N}$ & \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\end{center}
%
%

\section{Examples}
\label{sec:SP_Examples}
%
%
In this example we tell CFS++ to use the Bi-CGStab solver from the hypre
library together with an SPAI preconditioner:
%
%
\begin{verbatim}
<solver type="hypreBICGSTAB" precond="hypreSPAI">
  <matrix eqnNumbering="scalar" entry="double" storage="sparseNonSym"/>
  <hypreBICGSTAB>
    <tol>     1e-8 </tol>
    <maxIter> 1000 </maxIter>
    <logging>    2 </logging>
    <printLevel> 2 </printLevel>
  </hypreBICGSTAB>
  <hypreSPAI>
    <thresh> -0.4 </thresh>
    <levels>    6 </levels>
    <filter> -0.8 </filter>
    <symmetry>  0 </symmetry>
    <reuse>   yes </reuse>
  </hypreSPAI>
</solver>
\end{verbatim}
%
%
The following example specifies the native CG solver implemented in OLAS in
conjunction with a block numbering of the equations/unknowns. The system matrix
will be exported to the file ``piezo5.mtx'' and the right-hand side vector to 
he file ``piezo5.vec'':
\begin{verbatim}
<solver type="gmres">
  <matrix eqnNumbering="block"/>
  <exportLinSys baseName="piezo5"/>
  <gmres>
    <tol>1e-10</tol>
    <maxIter>2</maxIter>
    <logging>yes</logging>
  </gmres>
</solver>
\end{verbatim}

