\chapter{Howtos}

\section{Mounting Directories Using sshfs}

By  making use  of the  user file  system service  FUSE one  can  easily mount
dirctories from machines which are reachable  via SSH. The only thing to watch
out for, is to add the corresponding user to the fuse user group.

\subsection{Mounting HOME directory of laptop on a cluster node using sshfs}

Due  to the  fact,  that everybody  at  Klagenfurt uses  a  laptop to  perform
programming work,  most people  have local working  copies. In order  to avoid
copying of  these working  copies, it is  most convenient  to mount them  to a
directory  on  the cluster  servers  and to  do  compilation  to local  binary
directories on these machines.

The simplest  way of accomplishing this task  is to use sshfs.  By issuing the
follwoing command  on e.g. on sedici the  home directory of the  user simon on
the laptop mac will get mounted to the directory home\_mac on sedici.

\begin{verbatim}
sedici$ sshfs simon@mac: home_mac
\end{verbatim}

Please  make sure  to  do  the compilation  in  a local  directory  and not  a
subdirectory of home\_mac because not  all file system requirements are met to
do compilation there.

\subsection{Bridging Firewalls using SSH Tunnels}

If a  machine is not  directly reachable it  may become necessary to  bridge a
connection using  cascaded SSH Tunnels. In  the following scenario  we want to
mount the home directory of user  simon on mac at Klagenfurt (which is located
in a  VPN) on the machine  nc2 at Simetris GmbH  (which is also in  a VPN). In
order to do so we use the  intermediate machine lse81 at the LSE department in
Erlangen. The first  step is to connect  to lse81 from mac and  creating a SSH
tunnel which forwards the SSH port 22 on mac to port 22022 on lse81.

\begin{verbatim}
mac$ ssh -R 22022:mac:22 strieben@lse81.e-technik.uni-erlangen.de
\end{verbatim}

The next  step is  to forward the  port 22022  on lse81 to  port 22022  nc2 at
Simetris by issuing the following command on nc2.

\begin{verbatim}
nc2$ ssh -L 22022:localhost:22022 strieben@lse81.e-technik.uni-erlangen.de
\end{verbatim}

In a different shell  on nc2 one can now mount the home  directory on mac to a
directory on nc2 by issuing

\begin{verbatim}
nc2$ sshfs -p 22022 simon@localhost: home_mac
\end{verbatim}

\subsection{Unmounting a Share}

To unmount a sshfs share enter

\begin{verbatim}
$ fusermount -u home_mac
\end{verbatim}

\subsection{Mounting via SSH on other operating systems}
\begin{description}
\item[Linux]    For   most   distributions    there   are    binary   packages
  available.    Otherwise    SSHFS     can    be    compiled    from    source
  \url{http://fuse.sourceforge.net/sshfs.html}

\item[Mac  OS X] On  the Mac  one can  either install  MacFUSE and  SSHFS from
  \url{http://code.google.com/p/macfuse/wiki/MACFUSE_FS_SSHFS}      or     the
  commercial (40\$) solution ExpanDrive \url{http://www.expandrive.com/mac}.

\item[Windows]  On Windows  there exist  two  possible solutions.  There is  a
  similar     project     like     FUSE,     which     is     called     Dokan
  \url{http://dokan-dev.net/en}  and which is  freeware. It  is however  in an
  early  stage of  development  and  maybe very  slow.  The other  application
  ExpanDrive is commercial (40\$) \url{http://www.expandrive.com/windows}.
\end{description}

\subsection{Other FUSE File Systems}

\begin{description}
\item[CurlFtpFS] is for locally  mounting FTP directories 
  \url{http://curlftpfs.sourceforge.net/}

\item[NTFS-3G] read/write access to NTFS file systems on the Mac
  \url{http://www.tuxera.com/products/ntfs-open-source/}.

\item[davfs2]        is        for        mounting        WebDAV        shares
  \url{http://savannah.nongnu.org/projects/davfs2}.

\item[VMDK]  by  using  FUSE   the  vmware-mount  utility  from  Virtual  Disk
  Development Kit \url{http://www.vmware.com/support/developer/vddk/} on Linux
  can mount .vmdk hard disk images. The vmware-mount utility is also available
  on Windows.
\end{description}


\section{Debugging using Eclipse, Totalview, gdb}
\section{CFS++ Coding Style}
\section{Implementing a new PDE}
\section{Implementing a new solver}
\section{Implementing a new time stepping scheme}
\section{Implementing new readers/writers}
\section{Adding new Results}
\section{Video Howtos}
