\chapter{Software Infrastructure}

\section{Versioncontrolsystem Subversion}

Subversion is a  free/open source version control system,  which manages files
and  directories, and  the changes  made to  them, over  time. This  allows to
recover older versions and to have different developing branches \cite{svn}.

Git in combination with SVN: \url{http://git.or.cz/course/svn.html}

\section{CMake Build System}

CMake is  a family of platform  independent tools designed to  build, test and
package  software  \cite{cmake,  Schroeder2004,  Martin2007a}.  It  is  mainly
targeted towards C,  C++ or Fortran projects. Instead  of writing Makefiles or
autoconf/automake scripts  on Unix or  Visual Studio projects on  Windows, one
just  needs to maintain  platform independent  configuration files  from which
CMake will  generate the  previously mentioned projects  for the  native build
environments.

\section{CFSDEPS}

 When  developing a  general purpose  finite element  code a  lot  of external
libraries (e.g.   BLAS, LAPACK, ARPACK, Metis\ldots)  have to be  pulled in if
the wheel  is not  to be reinvented.  Many of  these libraries are  present on
modern-day Linux/Unix distributions.  If a library is however not available it
has to  be built  from source. If  more than  the standard GNU  compiler suite
should be supported it is  almost certain that incompatibilities with prebuilt
system  libraries  will   arise.  Therefore  we  decided  to   build  all  our
dependencies  in an  automated  script  based manner.  The  scripts fetch  the
sources from the web, unpack  them, perform checksum checks, patch, configure,
build   and   install   the   libraries.    This   way   we   obtain   maximum
platform-independence  while   having  a  well-documented   and  SVN-versioned
description of how to build external packages. The necessity for automatically
building external packages has also been recognized by the CMake developers in
the     meantime     and      is     supported     since     version     2.8.0
(cf. \cite{BuildExtPackages})

\section{CTest}

In order to  keep the quality of  a software project high it  is imparative to
perform manual and automated tests.  The necessary tests can be specified also
in the same  simple CMake syntax.  Every user should run  those tests before a
commit of  changes or additional software code.  This way it can  be made sure
that all the old features are still working at least on the user's platform.


\section{CDash}

This is an  open source, web-based software testing  server, which aggregates,
analyzes and displays the results of software testing processes submitted from
different clients \cite{cdash}.  We use it to gather information about nightly
builds  on various  platforms. This  way we  can track  down errors  which are
platform-dependent (e.g. 32 bit vs.  64 bit) more efficiently.


\section{Trac}

This  tool  is  an  enhanced  wiki  and issue  tracking  system  for  software
development projects. Therewith, we can discuss different issues, fix the next
development  steps   by  assigning   tickets,  browse  the   {\sf  Subversion}
repository, can look at the changes been made between different versions, etc.
\cite{trac}