\chapter{Quickstart Guide}

\section{List of external packages which need to be installed from distro}

The  list  of  required  packages  depends on  the  operating  system  and
especially for  Linux based  OSes on the  distribution.  We give  here the
commands for installing the required  packages from the command line.  The
package lists are provided  for some of the OSes most often  in use by the
CFS++ developers. If  you use a different distribution  you should be able
to figure out the required packages quite easily.

\subsection{openSUSE (11.2)}

\begin{verbatim} 
sudo zypper install subversion gcc gcc-c++ gcc-fortran automake autoconf \
                    cmake perl-base graphviz texlive-latex texlive-tex4ht \
                    python-pygments doxygen tcl-devel python-devel git-svn  
\end{verbatim}
\todo[inline]{Add packages for ccmake, cmake-gui and javac.}

\noindent If  some package  is not available  from the  standard repositories,
there   is  a   good  chance   to  find   it  at   openSUSE   software  search
\url{http://software.opensuse.org/search}   or  one   of  the   other  package
repositories like PackMan (\url{http://packman.links2linux.org}).

\noindent Checking which package provides an already installed file:
\begin{verbatim}
rpm -q --whatprovides /usr/bin/g++
\end{verbatim}

\subsection{Ubuntu (10.04)}

\begin{verbatim}
sudo apt-get install subversion gcc g++ gfortran automake autoconf cmake \
                     perl-base graphviz texlive-latex-base tex4ht \
                     python-pygments doxygen tcl-dev python-dev git-svn \
                     cmake-curses-gui cmake-qt-gui
\end{verbatim} 
\todo[inline]{Add package for javac.}

\begin{verbatim}
ln -s /usr/include/python2.6 /usr/include/python
\end{verbatim}

\noindent Checking which package provides an already installed file:
\begin{verbatim}
dpkg -S /usr/bin/g++
\end{verbatim}

\subsection{Mac OS X (10.5)}

\begin{itemize}
\item XCode - \url{http://developer.apple.com/technologies/tools/xcode.html}
\item MacPorts - \url{http://www.macports.org}
\item CMake - \url{http://www.cmake.org}
\item g++, tcl, etc. from MacPorts cf. \url{http://pc.freeshell.org/comp/macbook}
\end{itemize} 

\subsection{Windows (XP, Vista, 7)}

At the  moment we  only support the  following sub-projects on  Windows: NACS,
[CFS | NACS]DEPS.

\begin{itemize}
\item Visual Studio (Express) \url{http://www.microsoft.com/express/}
\item Platform SDK \url{http://en.wikipedia.org/wiki/Microsoft_Windows_SDK}
\item Subversion \url{http://tortoisesvn.tigris.org/}
\item UnxUtils - \url{http://unxutils.sourceforge.net/}
\item CMake - \url{http://www.cmake.org}
\item Perl \url{http://www.activestate.com/activeperl}
\item Python \url{http://www.activestate.com/activepython}
\item etc.
\end{itemize} 

Pre-installed   VBoxes  containg   all  required   packages  are   located  at
(Klagenfurt)

\noindent {\tt qnap:/share//AM\_Data/Software/NACS\_Windows\_Development\_VBOXes/[WINXP*\&nacs\_dev*\&local*]}.

\newpage
\section{Obtaining the Sources}

Once the software  needed for CFS++ development from  the distribution has
been installed,  the next  step is  to checkout the  source code  from the
Subversion  \cite{svn} repository. All  necessary steps  are given  in the
shell  script at  the end  of this  section. We  first create  a directory
structure for our  development resources. Next we obtain  the sources from
the server. Note,  that upon the first connection you  will have to accept
the server's  certificate permanently (p)  and decide, wether you  wish to
save your password locally.

\input{pygmentized/qs_obtain_sources.tex}
\attachfile[icon=PushPin,description={Shell script for obtaining CFS++ sources.},mimetype=text/plain]{attachments/qs_obtain_sources.txt}

Further information  about Subversion can be found  in Sec. \ref{sec:svn}.

A nice alternative, especially for working on a laptop, is the distributed
versioning system Git.  A basic howto for obtaining a  Git working copy of
the CFS++ sources, some basic commands and pointers to additional material
are given in Sec. \ref{sec:git}.

\section{Configuring the Build Tree and Building a Binaries}

We use the CMake build  system (cf. Sec. \ref{sec:cmake}) to configure our
source  tree  for building  on  a wide  range  of  platforms.  Unlike  the
well-known configure/make/make install methodology from Unix-like systems,
CMake  supports also  other platforms  like Windows.   It  generates build
descriptions for  the native development environments  (e.g.  Makefiles on
Unix and Visual  Studio projects on Windows) from  very simple input files
in  CMake syntax. Using  CMake, it  is also  possible to  do out-of-source
builds, which means,  that from a single source  tree multiple build trees
may generated (e.g. for debug or release executables). By doing dependency
analysis CMake can also generate Makefiles suitable for parallel builds.

In the following  script we create two empty build  directories. In one of
them we  will build a debug executable,  whereas in the other  one we will
build  a  release  executable.   Before  configuring we  set  our  desired
compilers by creating  new environment variables. After that  we change to
the  build  directories  and  configure   them.   This  step  may  take  a
considerable amount of  time for the first time  since a number libraries,
upon which  CFS++ depends, have to  built first. For this  purpose we also
have to specify  the location of CFSDEPS to CMake  and where the dowloaded
sources  and  prebuilt libraries  will  be stored  (CFS\_DEPS\_CACHE\_DIR,
cf. Sec. \ref{sec:cfsdeps}). We also  specify, that we want to use standard
Unix make files for building. The  configure step for the debug build will
be much faster, since the prebuilt libraries from the first step just need
to be unpacked into the debug directory.

\input{pygmentized/qs_configure_build}

\section{Generating a Project for Eclipse}
