\chapter{Mathematical Expressions}
\label{chap:MathExpr}
For many values (boundary conditions, load values) it is possible not
only to prescribe a constant value but also a complex mathematical
expression. CFS++ utilizes for this a customized version of the {\em
  muParser} library (\url{http://muparser.sourceforge.net/}).

\noindent
Therefore one can easily
\begin{itemize}
\item prescribe analytical given, space-dependent, boundary- and load
  conditions, like e.g. a parabolic flow profile or a linear varying load
\item use time and frequency dependent variables to incorporate time-/
  frequency-dependent values
\item generate an analytical time-function, like $sin(\omega t
  + \phi)$ without actually loading it from an external timedata file
  with discrete values
\end{itemize}

\noindent
Currently the following xml-elements allow the prescription of mathematical
expressions:
\begin{itemize}
\item \xel{dirichletInhom} for the \xat{value} and
\xat{phase} attribute
\item \xel{load} for the \xat{value} and \xat{phase}
attribute
\item \xel{regionLoad} in the mechanic PDE for the attribute
\xat{value}
\end{itemize}

\section{Built-in expressions and operators}
{\em muParser} already comes with an extensive set of pre-defined
expressions and and functions (see following two tables).

Note that it is currently not possible to define own variables within
the parser, as otherwise any occurrence of a not explicitly defined
variable would create a new one, initialized with zero.
\begin{table}[ht]
\begin{center}
 \caption{Built-in functions}
 \vspace{0.5cm}  
  \begin{tabular}{lll}
    \toprule
    name & \# of args. & description \\
    \midrule 
    \myverb#sin#     & 1 & sine function \\
    \myverb#cos#     & 1 & cosine function \\
    \myverb#tan#     & 1 & tangens function \\
    \myverb#asin#    & 1 & arcus sine function \\
    \myverb#acos#    & 1 & arcus cosine function \\
    \myverb#atan#    & 1 & arcus tangens function \\
    \myverb#sinh#    & 1 & hyperbolic sine function \\
    \myverb#cosh#    & 1 & hyperbolic cosine \\
    \myverb#tanh#    & 1 & hyperbolic tangens function \\
    \myverb#asinh#   & 1 & hyperbolic arcus sine function \\
    \myverb#acosh#   & 1 & hyperbolic arcus tangens function \\
    \myverb#atanh#   & 1 & hyperbolic arcur tangens function \\
    \myverb#log2#    & 1 & logarithm to the base 2 \\
    \myverb#log10#   & 1 & logarithm to the base 10 \\
    \myverb#log#     & 1 & logarithm to the base 10 \\
    \myverb#ln#      & 1 & logarithm to base e (2.71828...) \\
    \myverb#exp#     & 1 & e raised to the power of x \\
    \myverb#sqrt#    & 1 & square root of a value \\
    \myverb#sign#    & 1 & sign function -1 if $x<0$; 1 if $x>0$ \\
    \myverb#rint#    & 1 & round to nearest integer \\
    \myverb#abs#     & 1 & absolute value \\
    \myverb#if#      & 3 & if ... then ... else ... \\
    \myverb#min#     & var. & min of all arguments \\
    \myverb#max#     & var. & max of all arguments \\
    \myverb#sum#     & var. & sum of all arguments \\
    \myverb#avg#     & var. & mean value of all arguments \\
    \bottomrule
  \end{tabular}
\end{center}
\end{table}

\begin{table}[ht]
\begin{center}
 \caption{Built-in operators and constants}
 \vspace{0.5cm} 
  \begin{tabular}{lll}
    \toprule
    operator & \ meaning & priority \\
    \midrule
    \myverb#=#                  & assignment      & -1 \\
    \myverb#and#                & logical and      & 1 \\
    \myverb#or#                 & logical or       & 1 \\
    \myverb#xor#                & logical xor      & 1 \\
    \myverb#<=# or \verb#le# *  & less or equal    & 2 \\
    \myverb#>=# or \verb#ge# *  & greater or equal & 2 \\
    \myverb#!=#                 & not equal        & 2 \\
    \myverb#==#                 & equal            & 2 \\
    \myverb#># or \verb#gt# *   & greater than     & 2 \\
    \myverb#<# ot \verb#lt# *   & less than        & 2 \\
    \myverb#+#                  & addition         & 3 \\
    \myverb#-#                  & subtraction      & 3 \\
    \myverb#*#                  & multiplication   & 4 \\
    \myverb#/#                  & division         & 4 \\
    \myverb#^#                  & raise x to the power of y & 5 \\
    \myverb#pi#   or \verb#_pi# & $\pi = 3.1415926..$ & \\
    \myverb#_e#                 & Eulerian number $e = 2.7182818..$ \\
 	\bottomrule
  \end{tabular}
\end{center}
  \vspace*{0.5cm}
  {\footnotesize * Due to the restriction in the xml language definition no '$<$' or
    '$>$' signs are allowed in tokens. Therefore the latter form is the
    only one possible in .xml-files. If scripting is used, both forms are
    possible. }
\end{table}
\clearpage
\section{CFS custom functions}
\subsection{\texttt{sample1D}}
We can subscribe boundary conditions with values read in from sample files.
The notation is as follows:
\begin{Verbatim}[commandchars=@\[\]]
@PYaZ[<bcsAndLoads]@PYaZ[>]
  @PYaZ[<neumannInhom] @PYaQ[name=]@PYaB["transducer"] @PYaQ[value=]@PYaB["sample1D('x_amplitude.dat',x,2)"]
    @PYaQ[phase=]@PYaB["sample1D('x_phase.dat',x,2)"]@PYaZ[/>]
@PYaZ[</bcsAndLoads>]
\end{Verbatim}

Sample data (2$^{nd}$ column) is read in from the files x\_amplitude.dat and
x\_phase.dat. Please note the type of quotes used!!!!
\newline

\noindent
x specifies the direction, where the sample data is applied.
2 means cubic spline interpolation between sample values. 
Could also be 0=nearest neighbour, or 1=linear interpolation.

\subsection{\texttt{input}}
The \xel{input} element can be used to read in data on a whole region from
hdf5 files (of a previous simulation run). It takes as arguments the input ID
and the number of the multisequence step (which should be 1 in almost any
case). It is used in the following way:
\begin{Verbatim}[commandchars=@\[\]]
@PYaZ[<fileFormats]@PYaZ[>]
  @PYaZ[<input]@PYaZ[>]
    @PYaZ[<hdf5] @PYaQ[fileName=]@PYaB["filename.h5"] @PYaQ[id=]@PYaB["inhom"]@PYaZ[/>]
  @PYaZ[</input>]
  ...
@PYaZ[</fileFormats>]
...
@PYaZ[<pdename]@PYaZ[>]
  @PYaZ[<bcsAndLoads]@PYaZ[>]
    @PYaZ[<dirichletInhom] @PYaQ[name=]@PYaB["region1"] @PYaQ[dof=]@PYaB["x"] @PYaQ[value=]@PYaB["input('inhom',1)"]
                    @PYaQ[phase=]@PYaB["input('inhom',1)"]@PYaZ[/>]
    @PYaZ[<dirichletInhom] @PYaQ[name=]@PYaB["region2"] @PYaQ[dof=]@PYaB["y"] @PYaQ[value=]@PYaB["input('inhom',1)"]
                    @PYaQ[phase=]@PYaB["input('inhom',1)"]@PYaZ[/>]
  @PYaZ[</bcsAndLoads>]
@PYaZ[</pdename>]
\end{Verbatim}

The region name, quantity, and time/frequency step (if applicable) are
determined automatically and not passed as arguments. Therefore it is required
to use exactly same mesh as the previous simulation, i.e.\
interpolation onto a different mesh is not supported.

The \texttt{input} function is currently available only for inhomogeneous
Dirichlet boundary conditions, but may be supported for other BCs in the future.

\begin{table}[htb]
\begin{center}
\caption{CFS specific functions}
\vspace{0.5cm}
  \begin{tabular}{lll}
    \toprule
    name & \# of args. & description \\
    \midrule
    \texttt{input}      & 2 & read data for a whole region from a hdf5 file \\
    \texttt{sample1D}   & 3 & interpolation of boundary condition values
    from sample data files \\
    \bottomrule
  \end{tabular}
\end{center} 
\end{table}

\section{CFS custom variables}
In addition to the built-in variables, CFS++ registers additionally allow
some variables, which can be used in the mathematical expression (see
the following table).
\begin{table}[ht]
\begin{center}
 \caption{CFS specific variables}
 \vspace{0.5cm}
  \begin{tabular}{lll}
    \toprule
     name & value & occurrence \\
    \midrule
    \myverb#t#     & Current time value       & \xel{transient} analysis \\
    \myverb#f#     & Current frequency value  & \xel{harmonic} analysis \\
    \myverb$x y z$ & Current global coordinate value & boundary conditions
    e.g:load \\ \myverb#r phi z# & Current local coordinate value & regionLoad
    with cylindrical coordinate system \\
    \bottomrule
  \end{tabular}
\end{center}
\end{table}

\section{Examples}
If one wants to prescribe a linear space dependent load, the following can be
written:
\begin{Verbatim}[commandchars=@\[\]]
@PYaZ[<bcsAndLoads]@PYaZ[>]
  @PYaZ[<load] @PYaQ[name=]@PYaB["loadNode"] @PYaQ[value=]@PYaB["10 * x"] @PYaQ[phase=]@PYaB["cos(y)"]@PYaZ[/>]
@PYaZ[<bcsAndLoads]@PYaZ[>]
\end{Verbatim}

Here \myverb#x# is used to denote the x-coordinate of each node
contained in the node list.
\newline

\noindent
To define a variable with a sine of 1 kHz one can simply write:
\begin{Verbatim}[commandchars=@\[\]]
@PYaZ[<bcsAndLoads]@PYaZ[>]
  @PYaZ[<dirichletInhom] @PYaQ[name=]@PYaB["bcNode"] @PYaQ[value=]@PYaB["10 * sin(2 * pi * 1e3 * t)"]@PYaZ[/>]
@PYaZ[<bcsAndLoads]@PYaZ[>]
\end{Verbatim}

The variable \xatv{t} reflects the current time value in a transient
simulation.
Of couse one can also combine space- and time dependency:
\begin{Verbatim}[commandchars=@\[\]]
@PYaZ[<bcsAndLoads]@PYaZ[>]
  @PYaZ[<dirichletInhom] @PYaQ[name=]@PYaB["bcNode"] @PYaQ[value=]@PYaB["10 * x * sin(2 * pi * 1e3 * t)"]@PYaZ[/>]
@PYaZ[<bcsAndLoads]@PYaZ[>]
\end{Verbatim}

A ramp signal, which goes from $0 \ldots 1$ linearly and then
stays at 1 is written via the \xatv{if}-statement as follows:
\begin{Verbatim}[commandchars=@\[\]]
@PYaZ[<bcsAndLoads]@PYaZ[>]
  @PYaZ[<dirichletInhom] @PYaQ[name=]@PYaB["bcNode"] @PYaQ[value=]@PYaB["if(t lt 1,t,1.0)"]@PYaZ[/>]
@PYaZ[<bcsAndLoads]@PYaZ[>]
\end{Verbatim}


